---
title: "EXAMSTA5001"
output: word_document
date: "2024-06-10"
---


## Q1

```{r}
library(sp) 
library(sf) 
library(mapview)
library(geoR)

```


```{r}
data(s100)

#Quick summary of the data
summary(s100)


```

```{r}
#Visualise the locations of the data in the x-y plane
library(gstat)
library(geoR)
library(sp)

plot(s100)


```


```{r}
#Compute and plot a sample variogram.
bin1 <- variog(s100, uvec=seq(0,1,l=11))
plot(bin1)

```

```{r}
#Fit exponential and wave variograms to the data by using ordinary least squares.
#Hint: use the initial values (1,0.5) and zero nugget.

ols.ne <- variofit(bin1, ini = c(1,0.5), nugget=0, weights="equal", cov.model = "exponential")

ols.nw <- variofit(bin1, ini = c(1,0.5), nugget=0, weights="equal", cov.model = "wave")

```


```{r}
#Plot the fitted variograms against the sample variogram.


#For the exponential
plot(bin1)
lines(ols.ne, lwd = 2)
legend(200, 6500,legend=c("OLS"),lty=c(1,1),lwd=c(2,1), cex=0.7)
```

```{r}
#For the wave
plot(bin1)
lines(ols.nw, lwd = 2)
legend(200, 6500,legend=c("OLS"),lty=c(1,1),lwd=c(2,1), cex=0.7)
```

#Compare the fitted exponential and wave models. Which of these models do you think is better? Provide your justification.

#It seems that the exponential variogram is a better fit from the least squares method than the wave model.


```{r}
#Plot the empirical directional variogram for 60 degrees as a line plot.
vario60 <- variog(sic.100, max.dist = 300, direction=pi/3)
plot(vario60, type="l")
```

```{r}
#Plot a multidirectional variogram (computed for 0, 45, 90, and 135 degrees). What does this plot suggest?
plot(variog4(sic.100), legend = TRUE)

legend(200, 9000,legend=c("OLS"),lty=c(1,1),lwd=c(2,1), cex=0.5)
```



## Q2

```{r}
#Plot locations of the data (in black colour)  and the two points  (0.2, 0.2) and (1, 1) (in red colour) on the same plot.

plot(s100$coords,xlab="Coord X", ylab="Coord Y",xlim=c(0,2),ylim=c(0,
1.2))
loci <- matrix(c(0.2, 0.2), ncol=2) 
text(loci, as.character(1), col="red")

loci2 <- matrix(c(1, 1), ncol=2) 
text(loci2, as.character(2), col="red")


```

```{r}
#What are the predicted values by kriging at these points?

kc4 <- krige.conv(s100, locations = loci,krige = krige.control(obj.m = ols.nw)) 

kc4 

```

#The predicted value for the wave model is 0.7218152. 

```{r}
#For the second one

kc4_2 <- krige.conv(s100, locations = loci2,krige = krige.control(obj.m = ols.nw)) 

kc4_2 

```

#The predicted value for the second loci is 0.9872652


```{r}
#Perform kriging prediction on a grid covering the observation area,  display the result as an image.
pred.grid <- expand.grid(seq(0,2, l=101), seq(0,3, l=101)) 
 kc <- krige.conv(s100, loc = pred.grid,krige = krige.control(obj.m = ols.nw)) 
image(kc, loc = pred.grid) 


```

```{r}
#Cross-validate your model by using the leave-one-out cross-validation. Produce a bubble plot of the result.
a <- as.data.frame(s100$data)
s <- SpatialPointsDataFrame(s100$coords, a, 
proj4string=CRS(projargs=as.character(NA)), match.ID=TRUE)
v.fit <- as.vgm.variomodel(ols.nw)
cvLOOCV <- krige.cv(s100.data~1, s, v.fit, nfold=nrow(s))
bubble(cvLOOCV, "residual", main = "LOOCV residuals") 

```

#Explain the results in the plot.
#As large circles are approximately uniformly distributed over the region, the cross-validation analysis shows that there are no areas where the model's performance is substantially worse than in other regions. 

```{r}
#Cross-validate your model by using 5-fold cross-validation.
cv10 <- krige.cv(s100.data~1, s, v.fit, nfold=5) 
bubble(cv10, "residual", main = "log(zinc): 5-fold CV residuals") 

```

```{r}
#Optimize the monitoring network using the criterion of minimum mean kriging variances.  Which data point has the max value of mean kriging variance?

 m1 <- sapply(1:length(s), function(x) mean(krige(s100.data~1,s[-x,], s,v.fit)$var1.var)) 

 which(m1 == max(m1)) 

```

#Answer: 81

```{r}
#Plot the sorted values of the mean kriging variance.

plot(sort(m1)) 

```
```{r}
#Plot points with top and bottom 10%  of values of the mean kriging variance (use different point shapes for top and bottom points). Explain the plot.

layout(matrix(1:2, 1, 2)) 
plot(s100$coords) 
points(s100$coords[m1 < quantile(m1,.1),], pch=17) 
plot(s100$coords) 
points(s100$coords[m1 > quantile(m1,.9),], pch=16)


```




## Q3

```{r}
#Plot spatial locations of the data.
library(spatstat)
data(amacrine)
summary(amacrine)
```

```{r}
plot(amacrine)

```


```{r}
#Produce a contour plot for the estimated intensity.
den <- density(amacrine)
contour(den) 

```

```{r}
#Investigate the intensity of the data set. Do you think that the intensity is constant? Why?
lambda <- summary(amacrine)$intensity 
lambda
```
#Answer: 183.6113
#The intensity is constant when the data amacrine is stationary


```{r}
#Perform quadrat counting by dividing the data area into quadrats. Do you think that the point process is stationary? Explain your reasons.
quadratcount(amacrine, nx = 6, ny = 3)



```



```{r}
Q <- quadratcount(amacrine, nx = 6, ny = 3)
plot(amacrine, cex = 0.5, pch = "+")
plot(Q, add = TRUE, cex = 1)

```

#The preceding analysis demonstrates that the intensity of the spatial process is constant. Specifically, it decreases as the y coordinate. However, along the x coordinate, the intensity appears to remain relatively stable and can be assumed to be constant as well.

```{r}
#Separate the data into the sub-patterns of points of types "on" and "off" and plot their intensities.


#Do you think that the intensities are constant? Why? 
#For the off plot it appears to be constant cross the x and y coordinates. Similarly for the on plot as well.  
 plot(density(split(amacrine))) 

```


```{r}
#Plot the cross-type pair correlation function for "on" and "off" marks. Explain the plot and findings.

p <- pcfcross(amacrine, "off", "on") 
plot(p) 


#EXPLAINATION: The plot of the cross-type pair correlation function suggests that there is inhibition between "off" and "on" marks at very short distances and no inhibition at all other distances. 

```



```{r}
#Remove the marks from the point pattern and perform the spatial Kolmogorov-Smirnov test for the uniform distribution of the x coordinate. Summarize the findings

 testx <- cdf.test(amacrine, function(x, y) {x})
 plot(testx)
 legend(200, 9000,legend=c("observed","expected"),lty=c(1,1),lwd=c(2,1), cex=0.5)

```

```{r}
 cdf.test(amacrine, function(x, y) {x})

```
#The result shows that the p-value equals 0.973. Thus, the test does not reject the hypothesis of CSR. The plot displays the observed and expected distribution functions that also confirm it.Furthermore, While the Kolmogorov-Smirnov test does not reject complete spatial randomness along the x-axis, it fails to do so along the y-axis. Therefore, this pattern cannot be classified as completely spatially random.


```{r}
#Investigate the obtained point pattern using F and G functions. Explain the results.


#data<-data(amacrine) 
#plot(Fest(data)) 

```


```{r}
 
 #plot(Gest(mydata)) 


```
#Both functions suggest clustering behaviour as the theoretical curve G is under empirical estimators and vice versa for the function F.

## Q4

```{r}
#Write a code to simulate this Poisson process.
library(stpp)
data<-data(fmd)
str(fmd)
lbda1 <- function(x, y, t) {1000*exp(-4*y)*exp(-2*t)}

ipp <- rpp(lambda = lbda1)
plot(density(ipp$xyt[,3]))


```





```{r}
#Give a static display of the result.
#df1 <- data.frame( x = mydata$x, y = mydata$y, t = hours_diff ) 
#df1 <- df1[complete.cases(df1), ] 
#X1 <- as.3dpoints(df1) 
#str(X1) 
#plot(X1) 


```




```{r}
#Produce a two-panels plot showing spatial locations and cumulative times of events.
#animation(fmd,runtime=40,cex=0.5,s.region = northcumbria)
```