---
title: "MAST30025 Week 12 and 13 Lab R Code"
output: word_document
---


#REVISE THEM ON YOUR OWN TIME!
#Question 1: 
#Part a:Estimate the difference between treatment effects, and test if it is significantly different from 0.
```{r}
mydata = data.frame(response = c(7.5, 9.6, 8.4, 10.6, 9.9, 10.6,9.5, 9.7, 10.8, 11.9, 10.0, 12.9),treatment = factor(rep(c(1,2), c(6,6)))) #Ensure you convert these values as factors because we do not want R to assume it is at full rank!
summary(lm(response ~ treatment, mydata))
```
#The estimated difference between treatment effects is 1.3667. The p-value for the test of τ2 − τ1 = 0 (against a general alternative) is 0.0989, which is borderline significant (but insignificant at a 5% level).


#Part b:Now suppose that it is discovered that the response can be affected by the season, and that the data was collected over a period of six months, in the order given by the table. That is, a month was spent collecting each row of the table.
# We re-express the experiment by blocking: each month (row of the table) is considered one block, and we model the data as an additive two-factor model (the factors being the treatment and the block). Using this model, repeat your analysis. Is the estimate different? Is the p-value different?
```{r}
mydata$block = factor(rep(1:6, 2))
model = lm(response ~ treatment + block, mydata)
summary(model)
```

```{r}
anova(model)
```
#The estimated difference between treatment effects is still 1.3667. It has not changed be- cause the new design is a complete block design. The p-value for the test of τ2 − τ1 = 0 (against a general alternative) is now 0.0246, which is significant. The blocking has reduced the variability of our estimate and thus increased the power of our test.


#Question 4b:Give the design matrix X^A for a model with block and treatment effects (and an overall mean).
```{r}
XA = matrix(0, nrow = 12, ncol = 11)
XA[,1] = 1
XA[cbind(1:12,rep(1:6,each=2)+1)] = 1
XA[cbind(1:12,c(1,2,1,3,1,4,2,3,2,4,3,4)+7)] = 1
XA
```



#Question 4c: Using this model, estimate τ1 − τ2, the difference between the first two treatment effects, and its variance. Write the variance estimate as s2cT (XAT XA)cc for a suitable c.
```{r}
library(MASS)
library(Matrix)
y = c(1.245, 1.804, 2.468, 6.664, 5.573, -0.560,7.880, 10.469, 0.457, -3.621, -4.291, -9.384)
bA = ginv(t(XA) %*% XA) %*% t(XA) %*% y
cA = c(0, 0, 0, 0, 0, 0, 0, 1, -1, 0, 0)
n = 12
rA = rankMatrix(XA)[1]
rA

```

```{r}
s2A = sum((y - XA %*% bA)^2)/(n-rA)
s2A

```



```{r}
t(cA) %*% bA

```


```{r}
t(cA) %*% ginv(t(XA) %*% XA) %*% cA

```



```{r}
 s2A * t(cA) %*% ginv(t(XA) %*% XA) %*% cA

```



#Part d: Give the design matrix XB for a model with just treatment effects (and an overall mean).
```{r}
XB = matrix(0, nrow = 12, ncol = 5)
XB[,1] = 1
XB[cbind(1:12,c(1,2,1,3,1,4,2,3,2,4,3,4)+1)] = 1
XB

```



#Part e: Using this model, estimate τ1 − τ2, the difference between the first two treatment effects, and its variance. Write the variance estimate as s2cT (XBT XB)cc for a suitable c.
```{r}
bB = ginv(t(XB) %*% XB) %*% t(XB) %*% y
cB = c(0, 1, -1, 0, 0)
rB = rankMatrix(XB)[1]
rB

```

```{r}
s2B = sum((y - XB %*% bB)^2)/(n-rB)
s2B
```

```{r}
t(cB) %*% bB
```


```{r}
t(cB) %*% ginv(t(XB) %*% XB) %*% cB
```

```{r}
s2B * t(cB) %*% ginv(t(XB) %*% XB) %*% cB
```




#Part f: Show that when going from model A (BIBD) to model B (CRD) the term cT (XT X)cc decreases, but s2 increases markedly. What does this indicate?
See above. This indicates that the blocks are effective in their intended purpose (reducing variance).


#Part g: Is your estimate for τ1 − τ2 the same or different for the two models? Why?
The estimates are different. This is because in the BIBD, the blocks are not orthogonal to the treatments.




#Quesiton 6 
#Part a: Which design is a block design!
SOLUTION: The first design is a complete block design: each treatment appears exactly once in each block.

#Part b: Write down the design matrix for each design. Hence show that τ2 − τ1 is estimable in each case.
```{r}
Xa = matrix(0, 12, 8)
Xa[,1] = 1
Xa[1:3,2] = 1
Xa[4:6,3] = 1
Xa[7:9,4] = 1
Xa[10:12,5] = 1
Xa[,6] = c(1,0,0, 0,1,0, 1,0,0, 0,0,1)
Xa[,7] <- c(0,1,0, 1,0,0, 0,0,1, 0,1,0)
Xa[,8] <- c(0,0,1, 0,0,1, 0,1,0, 1,0,0)
Xb <- Xa
Xb[,6] = c(1,1,0, 0,0,0, 0,0,1, 1,0,0)
Xb[,7] = c(0,0,1, 1,1,0, 0,0,0, 0,1,0)
Xb[,8] = c(0,0,0, 0,0,1, 1,1,0, 0,0,1)
Xa
```



```{r}
Xb
```




```{r}
library(MASS)
t = c(0,0,0,0,0,-1,1,0)
round(t(t) %*% ginv(t(Xa) %*% Xa) %*% t(Xa) %*% Xa, 10)
```


```{r}
round(t(t) %*% ginv(t(Xb) %*% Xb) %*% t(Xb) %*% Xb, 10)
```


#Part c: For each design, in terms of the unknown error variance σ2, what is the variance of the estimator for τ2 − τ1, the difference between the first two treatment effects?
# Based on this, which design is better?

```{r}
t(t) %*% ginv(t(Xa) %*% Xa) %*% t
```



```{r}
t(t) %*% ginv(t(Xb) %*% Xb) %*% t
```

#The variance of the estimator is (1/2)σ2 for the first design and (2/3)σ2 for the second design. Clearly we prefer the CBD, as it gives the smaller variance.







































