---
title: "Workshop_7_STA5001"
output: word_document
date: "2024-04-22"
---


# Topic 1: Kriging
```{r}
library(geoR)
data(s100)
plot(s100$coords,xlim=c(0,1.2),ylim=c(0,1.2),xlab="CoordX",ylab="CoordY")
loci <- matrix(c(0.2,0.6,0.2,1.1,0.2,0.3,1.0,1.1),ncol=2)
text(loci,as.character(1:4),col="red")
polygon(x=c(0,1,1,0),y=c(0,0,1,1),lty=2)

```

## One can apply the commands below to perform the ordinary kriging by using
# the parameters estimated by ordinary least squares with nugget as.

```{r}
bin1<-variog(s100,uvec=seq(0,1,l=11))
ols.n<-variofit(bin1,ini=c(1,0.5),nugget=0.5,weights="equal",cov.model="exponential")
kc4<-krige.conv(s100,locations=loci,krige=krige.control(obj.m=ols.n))

```

## The predicted values at the four specified locations can be found as 

```{r}
kc4

```


#Topic 1 Lecture 10

```{r}
 pred.grid<- expand.grid(seq(0,1,l=51),seq(0,1,l=51))
 kc<-krige.conv(s100,loc=pred.grid,krige=krige.control(obj.m=ols.n))
 image(kc,loc=pred.grid,col=gray(seq(1,0.1,l=30)),xlab="CoordX",ylab="CoordY")

```

#The grey level indicates values of s100 at the corresponding locations.

#Topic 1: Slide 12

```{r}
library(gstat)
library(sp)
data(meuse)
data(meuse.grid)
coordinates(meuse)<-c("x","y")
coordinates(meuse.grid)<-~x+y
spplot(meuse,"zinc",do.log=T)
v <-variogram(log(zinc)~1,meuse)
v.fit<-fit.variogram(v,vgm(1,"Sph",800,1))

```


#Topic 1: Slide 14
```{r}
 lz.sk<-krige(log(zinc)~1,meuse,meuse.grid,v.fit,beta=5.9)
 lz.ok<-krige(log(zinc)~1,meuse,meuse.grid,v.fit)

```


```{r}
lz.sk

```

```{r}
class(lz.sk)

```


#Topic 1: Lecture 14

#For the universal kriging we fit a variogram model to residuals first:
```{r}
 vt<-variogram(log(zinc)~sqrt(dist),meuse)
 vt.fit<-fit.variogram(vt,vgm(1,"Exp",300,1))
 lz.uk<-krige(log(zinc)~sqrt(dist),meuse,meuse.grid,vt.fit)

```

```{r}
 lz.uk$var.pred.sk<-lz.sk$var1.pred
 lz.uk$var.var.sk<-lz.sk$var1.var
 lz.uk$var.pred.ok<-lz.ok$var1.pred
 lz.uk$var.var.ok<-lz.ok$var1.var
 spplot(lz.uk,c("var1.pred","var1.var","var.pred.ok","var.var.ok","var.pred.sk","var.var.sk"),layout=c(2,3))

```

#Topic 2: Model diagnostic 
#Slide 3 
```{r}
library(gstat)
data(meuse)
coordinates(meuse)<-c("x","y")
sel100<-sample(1:155,100)
m.model<-meuse[sel100,]
m.valid<-meuse[-sel100,]

```


#Slide 4.Then we fit a variogram model and compute the R^2 statistic of the residuals:

```{r}
v100.fit <- fit.variogram(variogram(log(zinc)~1, m.model),vgm(1, "Sph", 800, 1))
 m.valid.pr <- krige(log(zinc)~1, m.model,m.valid,v100.fit)
 resid.kr <- log(m.valid$zinc)- m.valid.pr$var1.pred
 summary(resid.kr)

```





```{r}
resid.mean <- log(m.valid$zinc)- mean(log(m.valid$zinc))
R2 <- 1- sum(resid.kr^2)/sum(resid.mean^2)
R2

```

#A map with cross validation residuals can be obtained by
```{r}
m.valid.pr$res <- resid.kr
bubble(m.valid.pr, "res")

```


#Topic 2: Slide 10
#Leave-one-outcrossvalidation (LOOCV) visits a data point,and
#predicts the value at that location by leaving out the observed #value, and proceeds with the next data point.
.
```{r}
v.fit <-vgm(.59,"Sph",874,.04)
cvLOOCV <-krige.cv(log(zinc)~1,meuse,v.fit,nfold=nrow(meuse))
bubble(cvLOOCV,"residual",main= "log(zinc):LOOCV residuals")
```



#N-fold cross validation makes a partition of a data set into N parts.
#For all observation in a part, predictions are made based on the remaining N −1 parts. This is repeated for each of the N parts.
```{r}
cv155 <- krige.cv(log(zinc)~1, meuse, v.fit, nfold=5)
bubble(cv155, "residual", main = "log(zinc): 5-fold CV residuals")
```

