---
title: "Workshop_5_STA5001"
output: word_document
date: "2024-04-08"
---

#Lecture 2. Slide 5.
#Sample variogram in R
```{r }

s100

```


```{r }
library(geoR)
data(s100)
cloud1<-variog(s100,option="cloud",max.dist=1)
cloud2<-variog(s100,option="cloud",estimator.type= "modulus",max.dist=1)
bin1<-variog(s100,uvec=seq(0,1,l=11))
bin2 <-variog(s100,uvec=seq(0,1,l=11),estimator.type="modulus")

```

#Lecture 2. Slide 6.
#To plot the obtained variogram, we 
# create a matrix of 2 x 2 plots:


```{r}
par(mfrow = c(2,2))
plot(cloud1, main = "classical estimator")
plot(cloud2, main = "modulus estimator")
plot(bin1, main = "classical estimator")
plot(bin2, main = "modulus estimator")

```

## Theoretical and sample variograms can be plotted and visually compared. FOr example, to plot a figure which shows the theoretical variogram model used to simulate the data s100 and the estimated variogram, type
```{r }
plot(bin1)
lines.variomodel(cov.model = "exp", cov.pars = c(1,0.3),nugget = 0, max.dist = 1, lwd = 3)
legend(0.4, 0.3, c("exponential model", "empirical"),lty=c(1,0), lwd = c(3,1), pch=c(NA,1))

```

#Directional variogram in R

```{r }
 vario60 <- variog(s100, max.dist = 1, direction=pi/3)
 plot(vario60)
 title(main=expression(paste("directional, angle =",60*degree)))

```

#For a quick computation of variograms in four directions the function var-iog4 can be used.

```{r }
vario.4 <- variog4(s100, max.dist = 1)
plot(vario.4, lwd=2)

```

