---
title: "STA5001 Workshop 3"
output: word_document
date: "2024-03-05"
---
#TOPIC 1: SPATIAL objects in package sf
#Lecture 1. Slide 4
```{r}
library(sf)
p1 <- st_multipoint(matrix(c(0,1/2,0,1/2),ncol=2))
p2 <- st_multipoint(matrix(c(2/3,1,2/3,1),ncol=2))

# Combine them into a simple feature collection:

sfc <- st_sfc(p1, p2)
str(sfc)

```

#Lecture 1. Slide 5



```{r}
d_sf <- st_sf(sfc, attribute = 1:2)
str(d_sf)

```

#Lecture 1.Slide 5



```{r}
plot(d_sf)

```
#Lecture 1. Slide 6

```{r}
outer <- matrix(c(0,0,10,0,10,10,0,10,0,0),ncol=2,byrow=TRUE)
hole1 <- matrix(c(1,1,1,2,2,2,2,1,1,1),ncol=2, byrow=TRUE)
hole2 <- matrix(c(5,5,5,6,6,6,6,5,5,5),ncol=2, byrow=TRUE)

#Then we combine them and transform into the st polygon object. 
pts <- list(outer,hole1,hole2)
pl <-st_polygon(pts)
plot(pl,col="grey",border="blue")

```
#Lecture 1. Slide 7.
```{r}
#NOTE: Ensure you follow the directory correctly
CRAN_df <- read.table("CRAN051001a.txt", header = TRUE)
str(CRAN_df)


```

#Lecture 1. Slide 8.
```{r}
CRAN_sf <- st_as_sf(CRAN_df, coords = c("long", "lat"))
str(CRAN_sf)

```
#We assign the WGS84 projection, which has the EPSG code 4326:
```{r}
st_crs(CRAN_sf)
st_crs(CRAN_sf) <- 4326
st_crs(CRAN_sf)

```

#Lecture 1. Slide 9
#Plotting sf objects.
```{r}
plot(st_geometry(CRAN_sf))
```

#Lecture 1. Slide 10
```{r}
CRAN_sf_Brazil = CRAN_sf[CRAN_sf$loc == "Brazil", ]

#MUST add existing plot from Slide 9
plot(st_geometry(CRAN_sf))
plot(st_geometry(CRAN_sf_Brazil),add=T,col="red",pch = 1)
```
#Lecture 1. Slide 11.
```{r}
plot(CRAN_sf)
```



#Lecture 1. Slide 12
```{r}
plot((CRAN_sf)[4])
```

#Lecture 1. Slide 13
#USING THE PACKAGE MAPVIEW
```{r}
#NOTE: Ensure to re-install 
library(mapview)
mapview(CRAN_sf, col.regions = "red", fgb = FALSE)

```


#Lecture 1. Slide 14 
#Conversion between sf and sp
```{r}
library(sp)
library(lattice)

#install stars to get the st_as_sf() to work.
library(stars)
data(meuse.grid)
coordinates(meuse.grid) <- ~x + y
m.sf <- st_as_sf(meuse.grid)
str(m.sf)

```
#Lecture 1. Slide 15
```{r}
plot(m.sf)

```
#Lecture 1. Slide 16
```{r}
CRAN_sp <- as(CRAN_sf, "Spatial")
str(CRAN_sp)


```

#TOPIC 2: USING GEOGRAPHIC MAPS IN R.
#lecture 2. Slide 2.
#Geographic maps in R.
```{r}
library(maps)
map("state", interior = FALSE)
map("state", boundary = FALSE, col = "gray", add = TRUE)


```

#Lecture 2. Slide 4
#GADM is a spatial database of the location of the world’s administrative areas (or administrative boundaries) for use in GIS and similar software.
```{r}
library(sp)
con<-gzcon(url("http://biogeo.ucdavis.edu/data/gadm2.8/rds/AUS_adm1.rds"))
data <- readRDS(con)

```

#Lecture 2. Slide 5
```{r}
library(RColorBrewer)
col <- rainbow(length(levels(data$NAME_1)))
data$NAME_1 = as.factor(data$NAME_1)
spplot(data, "NAME_1", col.regions=col,main="Australia Regions", colorkey = FALSE, lwd=.4)
col <- rainbow(length(levels(data$NAME_1)))
spplot(data, "NAME_1", col.regions=col,main="Australia Regions", colorkey = FALSE, lwd=.4)

```

```{r}
#Ensure you know where to keep the file in your directory

#Get the gadm41_AUS.gpkg" from this site.
#https://geodata.ucdavis.edu/gadm/gadm4.1/gpkg/gadm41_AUS.gpkg

data <- st_read("gadm41_AUS.gpkg", layer="ADM_ADM_1")

```


```{r}
str(data)

```

#Lecture 2. Slide 9
#Plotting sp/sf objects using world map
#using world cities csv.
```{r}
#Ensure you know where to keep the file in your directory
cities <- read.csv("worldcities.csv", header = TRUE)
str(cities)

```


#Lecture 2.Slide 10
```{r}

cities <-cities[complete.cases(cities), ]
sum(cities$population > 10000000)

```

```{r}

mcities <- cities[cities$population > 10000000, ]
mcities <- mcities[, c("lng","lat", "population")]
mcities

```



#Lecture 2. Slide 11
```{r}
mcities$population <- mcities$population/10000000
mcities_coor <- cbind(mcities$lng, mcities$lat)
row.names(mcities_coor) <- 1:nrow(mcities_coor)
row.names(mcities) <- 1:nrow(mcities)
str(mcities_coor)
llCRS <- CRS("+proj=longlat +ellps=WGS84")
mcities_sp <- SpatialPoints(mcities_coor, proj4string = llCRS)
summary(mcities_sp)

```
#Lecture 2. Slide 12
```{r}
mcities_spdf <- SpatialPointsDataFrame(mcities_coor,mcities,proj4string = llCRS, match.ID = TRUE)
summary(mcities_spdf)

```
```{r}

#OPTIONAL 
#library(rgrass)
#library(terra)
#library(sf)
#library(sp)


#writeVECT(mcities_spdf,"GRASSmcities", v.in.ogr_flags=c("o", "overwrite"), ignore.stderr=TRUE)


```

#Lecture 2. Slide 13
```{r}
library(maps)
```


```{r}
map("world", interior = FALSE)
map("world", boundary = FALSE, col = "gray", add = TRUE)
plot(mcities_spdf, pch = 16, col = "red", cex = (mcities_spdf$population), add = TRUE)

```


#Lecture 2. Slide 14
```{r}
mcities_sf <- st_as_sf(mcities_spdf)
str(mcities_sf)
```


```{r}
mapview(mcities_sf, col.regions = "red", cex = "population",fgb = FALSE)

```


#Lecture 2. Lecture 16
```{r}

lnstr_sfg2 <- st_linestring(as.matrix(mcities_sf[1:6,c("lng", "lat"), drop = TRUE]))
lnstr_sfc <- st_sfc(lnstr_sfg2, crs = 4326)
mapview(lnstr_sfc, alpha.regions = 0, color = "red",lwd = 2, fgb = FALSE)


```










