---
title: "CSE5DEV Assessment 2 Michael Le 21689299"
output:
  html_document:
    df_print: paged
date: "2023-08-08"
---

### FOR THIS ASSESSMENT I USED R VERSION 4.3.1 IN WINDOWS TO APPLY
### NESSARILY PACKAGES BELOW

```{r}
#Before starting the Assessment 
#1. Familise with the directories
#2. Understand what the packages do, you may need to change into the 
# proper version of R to ultise later in this assessment

#Packages for TASK 1
library(tidyverse)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)


#Packages for TASK 2
#Install the following packages below, Otherwise install the other packages 
#including some dependencies.Last Resort, update the right version of R.

library(DescTools)
library(tidyselect)


library(Hmisc)
library(PerformanceAnalytics)
library(corrplot)
#GGPUBR
#Go to this link https://cran.r-project.org/web/packages/ggplot2/index.html download 
#and unzip the file, I am using Windows for example (ensure to download which browser you are currently using)
#into the R directory, ensure to read which version of R you should be using
library(ggpubr)


```
#packages you have to manually install
```{r}
#GGPUBR
#Go to this link https://cran.r-project.org/web/packages/ggplot2/index.html download 
#and unzip the file, I am using Windows for example (ensure to download which browser you are currently using)
#into the R directory, ensure to read which version of R you should be using
library(ggpubr)

```

### DATASET FOR TASK 1: STUDENT MARKS
#TASK 1A and 1B
#Step 1. We can now determine the dataset from Movies. 

```{r}
#START TASK 1A and TASK 1B
dat_smarks = read.csv("C:/Users/mikel/OneDrive/Desktop/CSE5DEV Assessment 2 Code/Datasets/Studentmarks.csv",header = TRUE)
names(dat_smarks)
```
#Step 2. Print the structure of the whole data
```{r}
str(dat_smarks)
```


#Step 3 Print out the Student Marks dataset to understand them later when performing 
#further analysis. luckily we do not need to clean the data since we 
#have all values filled in all of the columns.
```{r}
dat_smarks
```
#Step 4. Check if there are any missing values within the dataset for Student Marks.
```{r}
#Luckily there is no missing values in the Student Mark Dataset. 
summary(dat_smarks)

```
#Step 5. Remove all row duplicate, since they do not impact on the overall 
#data
```{r}

dat_smarks <- distinct(dat_smarks)
dat_smarks 


```

# Step 6. We mutate and convert in Part A numeric,Part B integer and Part D integer 
# to compute the columns for ages for ages1, ages 2 and total marks respectively. 

```{r}
dat_smarks <- dat_smarks %>% mutate(
  dat_smarks, ages1 = as.numeric((Sys.Date() - as.Date(dob,format="%d/%m/%Y"))/365.25),
  ages2 = as.integer((Sys.Date() - as.Date(dob,format="%d/%m/%Y"))/365.25),
  total_marks = X2020 + X2021 + X2022)

dat_smarks 
#END OF TASK 1A and 1B

```

#TASK 1C
#Step 1. Renaming the columns to Marks_2020, Marks_2021 and Marks_2022 to 
#remember which year it was marked.

```{r}
#Start Task 1C
dat_smarks <- dat_smarks %>% rename(Marks_2020 = X2020,Marks_2021=X2021,
                                    Marks_2022 = X2022)
dat_smarks 
```
#Step 2. Applying the gather function to ensure it is easlier to categorise between
#Year and Mark.

```{r}
dat_smarks <- gather(dat_smarks,'Year','Mark',4:6)
dat_smarks 
```
#Step 3. Using the seperate function to ensure we want the Year value by itself.

```{r}
dat_smarks <- separate(dat_smarks,Year,c('Character','Year'),sep='_')
dat_smarks 
```
#Step 4. We want to remove the Character to avoid reduancy in the student dataset.

```{r}
dat_smarks <- dat_smarks %>% select(-Character)
dat_smarks 
```


#Step 5. Using the Scatterplot using geom point() grouped by Year for
#C. Create a scatter plot for Studentname versus Marks of all three years. 

```{r}
#END TASK 1C
ggplot (data = dat_smarks, mapping = aes(x = Studentname,y=Mark,col=Year)) + geom_point (alpha = .7, 
size = 3)  + labs(title = 'Scatterplot between Marks vs Students',subtitle = 'Grouped by Year', x= 'Student Name', y= 'Marks')
  
```
#TASK 1D
#Step 1.We want to filter the who got at least 200 marks in total and create a bar chart for students vs totalmarks in the descending order of their marks.


```{r}
#START TASK 1D
dat_smarks_atl200 <- dat_smarks %>% filter(total_marks >=200)
dat_smarks_atl200 <- dat_smarks_atl200 %>% arrange(desc(total_marks))
ggplot(dat_smarks_atl200, aes(x= reorder(Studentname,-total_marks),y=total_marks)) + geom_bar(stat="identity", color='red',fill='red') + labs(title = 'Bar chart of Students who scored 200 or more',subtitle = 'in descending Order', x= 'Student Name', y= 'Total Mark')
#END TASK 1D

```

### TASK 2: DATASET MOVIES

#Data visualisation section

#Step 1.Import data into RStudio
#In this case, were given the required datasets.
#MobliePhoneData, Movies and Studentmarks.

#TASK 2A
#Set the directory first, to locate your datasets 
```{r}
#Start Task 2A
setwd("C:/Users/mikel/OneDrive/Desktop/CSE5DEV Assessment 2 Code/Datasets")

```


#Step 2. Were are now performing some data wrangling, by 
#opening one of the datasets from the folder,
#For instance we use the 
#data for Movies
```{r}
dat_mov = read.csv("C:/Users/mikel/OneDrive/Desktop/CSE5DEV Assessment 2 Code/Datasets/Movies.csv",header = TRUE)

```



#Step 3. Check the name function to print the name of columns.
```{r}
names(dat_mov)
```



#Step 4. Check the head of the dataset for Movies for observation
```{r}
head(dat_mov)

```

#Step 5. Observe if there any missing values for the dataset for Movies.
```{r}
summary(dat_mov)

```
#Step 6. Count how many missing values overall movie dataset.
```{r}
#Counting the missing values for data-set movies
#there is 101 in total.
sum(is.na(dat_mov))


```
#Step 7. Missing values for all the columns we will compute or delete rows
#Only if it has impact on the overall Movie data
```{r}

#Total missing values for Reviews
sum(is.na(dat_mov$Reviews))

#Total missing values for Duration
sum(is.na(dat_mov$Duration))

#Total missing values for Actor_3_facebook_likes
sum(is.na(dat_mov$Actor_3_facebook_likes))

#Total missing values for Actor_1_facebook_likes
sum(is.na(dat_mov$Actor_1_facebook_likes))

#Total missing values for Facenumber_in_poster
sum(is.na(dat_mov$Facenumber_in_poster))

#Total missing values for Actor_2_facebook_likes
sum(is.na(dat_mov$Actor_2_facebook_likes))

#Total missing values for Aspect_ratio
sum(is.na(dat_mov$Aspect_ratio))

```

#Step 8. Remove all row duplicate, since they do not impact on the overall 
#data
```{r}

dat_mov <- distinct(dat_mov)
dat_mov


```
#STEP 9. If you decide to impute missing values 
#Step 9a, Take the mean() of all missing columns
#Step 9b, Take the median() of all missing columns
#Step 9c, Take the Mode() of all missing columns

#NOTE: If you decide not to impute the data if and only,
#Option 1. Remove all rows or columns consist missing values, Proceed to Step 10.
#Option 2. DO NOTHING Proceed to Step 12.

#Step 9a. Impute all the missing values by taking the means
#for each columns of missing data 
```{r}

dat_mov$Reviews[dat_mov$Reviews == mean(dat_mov$Reviews)] <- NA
dat_mov$Duration[dat_mov$Duration==mean(dat_mov$Duration)] <- NA 
dat_mov$Actor_3_facebook_likes[dat_mov$Actor_3_facebook_likes==mean(dat_mov$Actor_3_facebook_likes)] <- NA
dat_mov$Actor_1_facebook_likes[dat_mov$Actor_1_facebook_likes==mean(dat_mov$Actor_1_facebook_likes)] <- NA
dat_mov$Facenumber_in_poster[dat_mov$Facenumber_in_poster==mean(dat_mov$Facenumber_in_poster)] <- NA
dat_mov$Actor_2_facebook_likes[dat_mov$Actor_2_facebook_likes==mean(dat_mov$Actor_2_facebook_likes)] <- NA
dat_mov$Aspect_ratio[dat_mov$Aspect_ratio==mean(dat_mov$Aspect_ratio)] <- NA


```

#Step 9b. Impute all the missing values by taking the medians
#for each columns of missing data (CHOSEN THIS ONE)
```{r}

dat_mov$Reviews[dat_mov$Reviews == median(dat_mov$Reviews)] <- NA
dat_mov$Duration[dat_mov$Duration==median(dat_mov$dat_mov$Duration)] <- NA 
dat_mov$Actor_3_facebook_likes[dat_mov$Actor_3_facebook_likes==median(dat_mov$Actor_3_facebook_likes)] <- NA
dat_mov$Actor_1_facebook_likes[dat_mov$Actor_1_facebook_likes==median(dat_mov$Actor_1_facebook_likes)] <- NA
dat_mov$Facenumber_in_poster[dat_mov$Facenumber_in_poster==median(dat_mov$Facenumber_in_poster)] <- NA
dat_mov$Actor_2_facebook_likes[dat_mov$Actor_2_facebook_likes==median(dat_mov$Actor_2_facebook_likes)] <- NA
dat_mov$Aspect_ratio[dat_mov$Aspect_ratio==median(dat_mov$Aspect_ratio)] <- NA


```


#Step 9c. Impute all the missing values by taking the modes
#for each columns of missing data
```{r}

dat_mov$Reviews[dat_mov$Reviews == Mode(dat_mov$Reviews)] <- NA
dat_mov$Duration[dat_mov$Duration==Mode(dat_mov$Duration)] <- NA 
dat_mov$Actor_3_facebook_likes[dat_mov$Actor_3_facebook_likes==Mode(dat_mov$Actor_3_facebook_likes)] <- NA
dat_mov$Actor_1_facebook_likes[dat_mov$Actor_1_facebook_likes==Mode(dat_mov$Actor_1_facebook_likes)] <- NA
dat_mov$Facenumber_in_poster[dat_mov$Facenumber_in_poster==Mode(dat_mov$Facenumber_in_poster)] <- NA
dat_mov$Actor_2_facebook_likes[dat_mov$Actor_2_facebook_likes==Mode(dat_mov$Actor_2_facebook_likes)] <- NA
dat_mov$Aspect_ratio[dat_mov$Aspect_ratio==Mode(dat_mov$Aspect_ratio)] <- NA


```



#Step 10. We then proceed, if we confident with our new filled in Movie Dataset to omit all the 
#the rows. 
```{r}

dat_mov <- na.omit(dat_mov)

```


#Step 11. Check if there is any more missing values and check the data types in the new updated data
#for movies. 

```{r}
#End of TASK 2A
summary(dat_mov)
str(dat_mov)

```

#TASK 2B.
#Step 1.  Based on the Dataset, calculate “Profit” and determine the relationship between 
#“Profit” and other variables (e.g. IMDB score). Hint: Profit = Gross – Budget. Use line 
#plot or scatter plot to find the relationship. Present a summary of your findings 

```{r}
#Starting TASK 2B.
dat_mov <- dat_mov %>% mutate(dat_mov,Profit =Gross-Budget)

```
#Step 2. 
#Plot all the nessarily ggplots scatterplots for profit vs (Imdb score, reviews,votes and Movie facebook likes) 

```{r}

#1.ggplot between Imdb_score vs. Profit
g1 <- ggplot(data = dat_mov, mapping = aes(x = Imdb_score ,y=Profit)) + geom_point (alpha = .7, size = 3) + labs(title = 'Imdb Score vs Profit',x='Imdb Score',y='Profit') 

#2.ggplot between Reviews vs. Profit
g2<-ggplot(data = dat_mov, mapping = aes(x = Reviews,y=Profit)) + geom_point (alpha = .7, size = 3) + labs(title = 'Reviews vs Profit',x='Reviews',y='Profit')  

#3.ggplot between Votes vs. Profit
g3<-ggplot(data = dat_mov, mapping = aes(x = Votes,y=Profit)) + geom_point (alpha = .7, size = 3) + labs(title = 'Votes vs Profit',x='Reviews',y='Profit') 

#4.ggplot between Movie Facebook Likes vs. Profit
g4<-ggplot(data = dat_mov, mapping = aes(x = Movie_facebook_likes,y=Profit)) + geom_point (alpha = .7, size = 3) + labs(title = ' Movie facebook likes vs Profit',x='Movie facebook likes',y='Profit')


```


#Step 3. Keep all ggplots into 1 Figure.
```{r}

#End of Part B
figure <- ggarrange(g1,g2,g3,g4,nrow=2,align="h",widths = c(1,1,1),ncol=2,heights=c(1,1,1))
figure

```


#TASK 2C
#Step 1. Calculate the correlation between the relevant variable(s) used in the Dataset. 
#Present a summary of your findings 

```{r}

str(dat_mov)

```
#Step 2. Calculate the correlation between the relevant variable(s) used in the Dataset. 
#Present a summary of your findings 

```{r}
#Start of Part C
#To start were only interest in columns that has numeric 
#values in the Movie Dataset

dat_mov <- dat_mov %>% select(-Color) 
dat_mov <- dat_mov %>% select(-Director) 
dat_mov <- dat_mov %>% select(-Actor_2_name) 
dat_mov <- dat_mov %>% select(-Genre)  
dat_mov <- dat_mov %>% select(-Actor_1_name) 
dat_mov <- dat_mov %>% select(-Title) 
dat_mov <- dat_mov %>% select(-Actor_3_name) 
dat_mov <- dat_mov %>% select(-Plot_keywords)
dat_mov <- dat_mov %>% select(-Movie_imdb_link) 
dat_mov <- dat_mov %>% select(-Language)
dat_mov <- dat_mov %>% select(-Content_rating)
dat_mov <- dat_mov %>% select(-Gross)
dat_mov <- dat_mov %>% select(-Reviews)
dat_mov <- dat_mov %>% select(-Duration)
dat_mov <- dat_mov %>% select(-Director_facebook_likes)
dat_mov <- dat_mov %>% select(-Actor_3_facebook_likes)
dat_mov <- dat_mov %>% select(-Actor_1_facebook_likes)
dat_mov <- dat_mov %>% select(-Votes)
dat_mov <- dat_mov %>% select(-Cast_total_facebook_likes)
dat_mov <- dat_mov %>% select(-Facenumber_in_poster)
dat_mov <- dat_mov %>% select(-Year )
dat_mov <- dat_mov %>% select(-Actor_2_facebook_likes)
dat_mov <- dat_mov %>% select(-Movie_facebook_likes)

#dat_mov <- dat_mov %>% select(-Budget)

names(dat_mov)
summary(dat_mov)




```
#Step 3. Check if there is any numeric values? 
```{r}
str(dat_mov)

```

#Step 4a. Visualise the numeric variables between Profit vs. other variables using scatterplots
#Using ggscatter()
```{r}
#Plot 1.Imdb score vs Profit
g5 <- ggscatter(dat_mov,x="Imdb_score",y="Profit",add="reg.line",cont.int=TRUE,xlab="Imdb Score",ylab="Profit")

#Plot 2.Aspect Ratio vs Profit
g6 <- ggscatter(dat_mov,x="Aspect_ratio",y="Profit",add="reg.line",cont.int=TRUE,xlab="Aspect ratio",ylab="Profit")

#Plot 3.Budget vs Profit
g7 <- ggscatter(dat_mov,x="Budget",y="Profit",add="reg.line",cont.int=TRUE,xlab="Budget ",ylab="Profit")

#Visualise Plots 1 and 2. 
figure_1 <- ggarrange(g5,g6,nrow=1,ncol=2,widths = c(1,1),align="v",heights=c(1,1))
figure_1

```
#Step 4b.Visualise Plot 3 Only
```{r}
#Visualise 
g7

```

#Step 5a. Implement Pearson correlation, Spearman correlation and Kendall correlation
#On all the scatterplots demonstrated in Steps 4a and 4b.
```{r}
#Plot 1.Imdb score vs Profit
g5 <- ggscatter(dat_mov,x="Imdb_score",y="Profit",add="reg.line",cont.int=TRUE,cor.coef=TRUE,cpr.method="pearson",xlab="Imdb Score",ylab="Profit",color="yellow")

#Plot 2.Aspect Ratio vs Profit
g6 <- ggscatter(dat_mov,x="Aspect_ratio",y="Profit",add="reg.line",cont.int=TRUE,cor.coef=TRUE,cpr.method="pearson",xlab="Aspect ratio",ylab="Profit",color="red")

#Plot 3.Budget vs Profit
g7 <- ggscatter(dat_mov,x="Budget",y="Profit",add="reg.line",cont.int=TRUE,cor.coef=TRUE,cpr.method="pearson",xlab="Budget ",ylab="Profit",color="lightblue")

#Visualise Plots 1 and 2. 
figure_1 <- ggarrange(g5,g6,nrow=1,ncol=2,widths = c(1,1),align="v",heights=c(1,1))
figure_1

```


#Step 5b.Same thing for plot 3.
```{r}

g7

```
#Step 6. Apply the Shapiro-Wilk Test

#Check the normality of Profit and Budget,Imdb Score and Aspect Ratio.
```{r}
shapiro.test(dat_mov$Profit)

```

```{r}
shapiro.test(dat_mov$Budget)

```


```{r}
shapiro.test(dat_mov$Imdb_score)

```

```{r}
shapiro.test(dat_mov$Aspect_ratio)

```
#From the results, we can see that p-values are less than the significance level 
#0.05. This means that H1 does not get rejected. Hence, we can assume
#Profit and other numeric values are not normally distributed.

#Step 7. Apply the Pearson correlation between Profit and other variables. 
```{r}
#1. Test Between Profit and Budget
res_p1 <- cor.test(dat_mov$Profit, dat_mov$Budget, method = "pearson")
res_p1

```

```{r}
#2. Test Between Profit and Imdb_score
res_p2 <- cor.test(dat_mov$Profit, dat_mov$Imdb_score, method = "pearson")
res_p2

```


```{r}
#3. Test Between Profit and Aspect Ratio
res_p3 <- cor.test(dat_mov$Profit, dat_mov$Aspect_ratio, method = "pearson")
res_p3

```
#Conclusion, Budget has a negative correlation compared to other variables where it has no correlation
#since we can conclude that they are independant.

#Step 8. Apply Spearman rank correlation coefficient between Profit and other variables.
```{r}
#1. Test Between Profit and Budget
res_s1 <- cor.test(dat_mov$Profit, dat_mov$Budget, method = "spearman")
res_s1

```

```{r}
#2. Test Between Profit and Imdb_score
res_s2 <- cor.test(dat_mov$Profit, dat_mov$Imdb_score, method = "spearman")
res_s2

```

```{r}
#3. Test Between Profit and Aspect Ratio
res_s3 <- cor.test(dat_mov$Profit, dat_mov$Aspect_ratio, method = "spearman")
res_s3

```

#Step 9. Lastly the Kendall rank correlation test
```{r}
#1. Test Between Profit and Budget
res_k1 <- cor.test(dat_mov$Profit, dat_mov$Budget, method = "kendall")
res_k1

```

```{r}
#2. Test Between Profit and Imdb_score
res_k2 <- cor.test(dat_mov$Profit, dat_mov$Imdb_score, method = "kendall")
res_k2

```

```{r}
#3. Test Between Profit and Aspect Ratio
res_k3 <- cor.test(dat_mov$Profit, dat_mov$Aspect_ratio, method = "kendall")
res_k3

#END OF TASK 2C

```
#END TASK 2C


#TASK 2D
#Step 1. Compute the correlation matrix here Profit is the dependant variable (y)
```{r}

#Swap the columns putting the column variable profit in the front.
dat_mov <- dat_mov[c("Profit","Budget","Imdb_score","Aspect_ratio")]
M <-cor(dat_mov)
corrplot(M,method="circle", addCoef.col = 1,number.cex = 0.5, tl.cex = 0.5)

#END TASK 2D

```

#COMMENT ON TASK 2D:
#Since we choosen Option 9b (after clearing duplicate rows and inputting only the means
#for each column). The Strong Negative Correlation that occur between Profit and Budget with 
#the correlation of -0.95. For Profit and imdb Score there is very weak correlation with the correlation of 0.04.
#Lastly, for Profit vs. Aspect Ratio has no correlation or imply a very weak positive relationship. 
