---
title: "Week 2 HarvardX Statistics and R"
output: word_document
---

## Random Variables Excercises 





```{r}
library(downloader) 
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/femaleControlsPopulation.csv"
filename <- basename(url)
download(url, destfile=filename)
x <- unlist( read.csv(filename) )

```

## Random number generator (RNG)

```{r}
RNGkind("Mersenne-Twister", "Inversion", "Rejection")

```


## Question 1
#What is the average of these weights?

```{r}
mean(x)

```
#Set the seed to 1.

```{r}
set.seed(1)

```


## Question 2
#Take a random sample of size 5. What is the absolute value (use abs()) of the difference between the average of the sample and the average of all the values?

```{r}
control = sample(x,5)

#Another 5 just in case.
treatment = sample(x,5)

abs(mean(treatment)-mean(control))

```

#Actual solution

```{r}
 set.seed(1)
X = sample(x,5)
abs(mean(X) - mean(x))

```

## Question 3
#After setting the seed at 5, set.seed(5), take a random sample of size 5. What is the absolute value of the difference between the average of the sample and the average of all the values?


```{r}
#Actual solution
set.seed(5)
X = sample(x,5)
abs(mean(X) - mean(x))
```
## Question 4
# Why are the answers from 2 and 3 different?

Answer: Because the average of the samples is a random variable. 











