---
title: "Untitled"
output: word_document
---

```{r}
library(downloader) 
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/femaleMiceWeights.csv"
filename <- "femaleMiceWeights.csv" 
download(url, destfile=filename)
```
#Exercise #1
#0 points possible (ungraded)
#Please note that this question does not count toward your grade, but it is important to make sure that you are using an appropriate version of R.

#What version of R are you using?

#3.6.3



#Exercise #2
#1/1 point (graded)
#Create a numeric vector containing the numbers 2.23, 3.45, 1.87, 2.11, 7.33, 18.34, 19.23.

What is the average of these numbers?
7.794286

#Exercise #3
1/1 point (graded)
Use a for loop to determine the value of sum i = 1 to 25 i^2

5525

#Exercise #4
1/1 point (graded)
The cars dataset is available in base R. You can type cars to see it. Use the class() function to determine what type of object is cars.

data.frame

#Exercise #5
1/1 point (graded)
How many rows does the cars object have?

50

#Exercise #6
1/1 point (graded)
What is the name of the second column of cars?

dist

#Exercise #7
1/1 point (graded)
The simplest way to extract the columns of a matrix or data.frame is using [. For example you can access the second column with cars[,2].

What is the average distance traveled in this dataset?
42.98

#Exercise #8
1/1 point (graded)
Familiarize yourself with the which() function. Which row of cars has a a distance of 85?

50



## Getting Started Exercises #1

```{r}
this = read.csv(filename)
this
```

##  Getting Started Exercises #2

```{r}

this[12,2]

```

##  Getting Started Exercises #3

```{r}

this$Bodyweight[11]

```


##  Getting Started Exercises #4

```{r}

weights <- this$Bodyweight
length(weights)
```



##  Getting Started Exercises #5

```{r}

View(this) 
weights <- this$Bodyweight
mean(weights[13:24])

```


##  Getting Started Exercises #6

```{r}
sample(this$Bodyweight[13:24])
set.seed(1)

```

```{r}
      set.seed(1)
i <- sample(13:24, 1)
this$Bodyweight[i]

```



