---
title: "part3_code"
output: word_document
date: "2023-10-17"
---

```{r}
bikes <- read.csv("C:/Users/mikel/OneDrive/Desktop/Assessment_2_STM4PSD/bikes.csv",header=TRUE)
bikes

```

## OBJECTIVES
## Only confidence intervals are needed. 

#The confidence intervals must use the form we look at this week (Week 5). Interpretations are the same across Weeks 4 and 5, but the calculations in Week 5 are more precise. However, the functions that do these computations will do most of the work automatically.

#Basically, you can use the t.test and prop.test functions to get all the required output - your task is to know when and how to use which variants of these functions. 


```{r}
#The average number of casual users, registered users, and total users per day
avg_casual <- mean(bikes$casual)
avg_reg <- mean(bikes$registered)
avg_total <- mean(bikes$casual+bikes$registered)

#Proportion of the registered users among all users
prop_reg <- sum(bikes$registered)/(sum(bikes$registered)+sum(bikes$casual))


```

# Differences in usage by casual users versus registered users.
```{r}



```

# Whether there are yearly differences in bikeshare usage.



#OUR GOAL IS TO PERFORM A LOT OF CONFIDENCE INTERVALS

#I.e. There are four types of Confidence Intervals 

#Single Mean
#Proportion
#Difference in two means 
#Unpaired Difference in two means


#Single Mean
```{r}



```










