---
title: "MAST30025 Week 8 Lab Q5 R Code"
output:
  pdf_document: default
  word_document: default
---


#Question 5

#Part a: Use the read.csv function to read the data. Then convert the type component into a factor
```{r}
setwd("~/Desktop/UNIMELB S1 2021 (Currently)/MAST30025/Tutorials /Tutorials/Rfile/data")
filters = read.csv("filters.csv")

#More to add how to convert into a factor! #Slide 63 Less than full rank model!
filters$type = factor(filters$type)
```


#Part b: Construct X and y matrices for this linear model!
```{r}
#Attempt 1

#y matrix
y = filters$life

#X matrix
n = dim(filters)
k = length(levels(filters$type))
X = matrix(0,n,k+1)
#Add all ones in the first column
X[,1] = 1
#Add all ones for each tao factor

#2nd column
X[filters$type==1,2] = 1
#3rd column
X[filters$type==2,3] = 1
#4th column
X[filters$type==3,4] = 1
#5th column
X[filters$type==4,5] = 1
#6th column
X[filters$type==5,6] = 1
```



```{r}
y

```

```{r}
X
```
```{r}
rankMatrix(X)[1]
```



```{r}

#Actual X matrix should works out exactly above
X1 = matrix(0,30,6)
X1[,1] = 1
for (i in 1:5){X1[filters$type==i,i+1] = 1}
X1
```


#Part c: Using the algorithm given in the lecture slides, find a conditional inverse for t(X)%*%X!
```{r}
#Actual X matrix should works out exactly above
t(X)%*%X
```



```{r}
#finding the rank in t(X)%*%X which is 5 
#Using Theorem 6.2 it is a 5 by 5 matrix and it has to be full rank and it is invertiable 
M = 6*diag(5)
M
```


```{r}

#finding the inverse
U = solve(M)

#then getting the conditional inverse
CI = data.frame(U)
R = rbind(c(0,0,0,0,0,0),CI)
C = cbind(X0= 0, R)
C
```

```{r}
#If you want it to convert into a matrix form!
#finding the inverse
XtXc = matrix(0,6,6)
XtXc[2:6,2:6] = solve((t(X)%*%X)[2:6,2:6])
XtXc
```

#Part d: Use ginv to find another conditional inverse for t(X)%*%X
```{r}
#Finding another conditional inverse (Open the MASS Package)
XtXc2 = ginv(t(X)%*%X)
XtXc2
```

#Part e: Verify that X(t(X)%*%X)^c%*%t(X) is the same for your two conditional inverses
```{r}
#Attempt 1
#Case 1 from Part c!
X%*%XtXc%*%t(X)

```


```{r}
#Attempt 1
#Case 2 from Part c!
X%*%XtXc2%*%t(X)

```

```{r}
#Actual Answer
sum((X%*%XtXc%*%t(X)-X%*%XtXc2%*%t(X))^2)

#Calculate the difference, in this case the difference here is very little. Safely assume that these computations are the same!

```


#Part f: Find two solutions for the normal equations
```{r}
#Attempt 1 for the first conditional inverse
b1 = XtXc%*%t(X)%*%y
b1
```


```{r}
#Attempt 1 for the second conditional inverse
b2 = XtXc2%*%t(X)%*%y
b2
```

#Part g: Express one of your solutions in terms of the other.
```{r}
#From slide 72/137 Less than the full rank model
#Sorry 1 step at a time
b2 = diag(6) - XtXc%*%t(X)%*%X
b2

```




#Part h: Write down a form for all solutions to the normal equations
```{r}
#ASK ABOUT IT, #All solutions to the normal equation are of the form b + (I-(t(X)%*%X)^c%*%t(X)%*%X)%*%z for arbitary z!

```












