---
title: "ASSIGNMENT 2 2009 MAST30025"
output: word_document
---

## Question 1:
#Part a: Calculate the least squares estimator b
```{r}
y = c(6.5,5.9,8.0,9.0,10.0,10.8)
X = matrix(c(rep(1,6),9,5.5,9,9.8,14.5,8,10,9,12,11,12,13.7,4,7,5,6.2,5.8,3.9),6,4)
b = solve(t(X)%*%X,t(X)%*%y)
b #estimates of the parameters
```

#Part b:In this year, the unemployment rate is 7% and the interest rate is 8%. What price should a dealer put on a car to expect to sell 7500 cars?

```{r}
n=6
p=4
e = y - X%*%b
SSRes = sum(e^2)
s2 = SSRes/(n-p)
s2 #Sample variance 
```


```{r}
(7.5 - b[1] - 7 * b[3] - 8 * b[4])/b[2]

```

#Part c: Find a 95% confidence interval for the average number of $10, 000 cars sold in a year which has unemployment rate 4.5% and interest rate 7%.
```{r}
xst = c(1,10,4.5,7)
ta = qt(0.975,df = 6-4)
xst%*%b+c(-1,1)*ta*s*sqrt(t(xst)%*%solve(t(X)%*%X)%*%xst)

```


#Question 2 Part a: Test for model adequacy.


#Attempt 1 (Correct!)

```{r}
SSReg = t(y)%*%X%*%b
SSReg
```



```{r}
SSTotal = sum(y^2)
SSTotal
```

```{r}
SSRes = SSTotal - SSReg
SSRes
```


```{r}
MSReg = SSReg/4
MSReg
```


#SSReg>SSRes


```{r}
MSRes = SSRes/2
MSRes
```


```{r}
Fstat = MSReg/MSRes
Fstat
```

```{r}
qf(0.95,4,2)
```

```{r}
pf(Fstat,4,2,lower.tail = FALSE)
```
#Model clearly relevant or adequate!


#Question 2 Part b: We want to find the best model for this data. Starting with an empty model, perform tests at the 5% level to determine if each parameter should be added to the model, in the order intercept,cost,unemployment,interest rate. If a parameter should be added, add it for the remaining test; otherwise, leave it out. Using this approach, what is the best model?


```{r}
CarsSold = c(6.5,5.9,8.0,9.0,10.0,10.8)
cost = c(9,5.5,9,9.8,14.5,8)
UnemploymentRate = c(10,9,12,11,12,13.7)
InterestRate = c(4,7,5,6.2,5.8,3.9)
demandcars.data = data.frame(CarsSold,cost,UnemploymentRate,InterestRate)
damandcarsdata = str(demandcars.data)
```


```{r}
basemodel11 = lm(CarsSold~1 , data = damandcarsdata) 
```

#Attempt 1 (Correct): But using a different method
```{r}
add1(basemodel11,scope = ~. + cost+UnemploymentRate+InterestRate, test = "F")
```

```{r}
basemodel12 = lm(CarsSold~UnemploymentRate, data = damandcarsdata) 
add1(basemodel12,scope = ~. + cost+InterestRate, test = "F")
```

#We use the variable, unemployment Rate in our final model!
#Off course the intercept!

#NOTE: Have to check the solution again!

#Part c: 
#Simultaneously test the hypotheses that the unemploy- ment rate is not relevant to the cars sold and that the parameters corresponding to cost and interest rate are equal!

#NOTE: Refer from Slide 32, Linear Statistical Models Inference for the full rank model!
```{r}
C <- matrix(c(0, 0, 0, 1, 1, 0, 0, -1), 2, 4)
Fstat <- (t(C %*% b) %*% solve(C %*% solve(t(X) %*% X) %*% t(C)) %*%C %*% b/2)/(SSRes/(6 - 4))
Fstat
```

```{r}
pf(Fstat,2,6-4,lower.tail = FALSE)
```
#Can not reject the null hypothesis!!

