---
title: "ASSIGNMENT 2 MAST30025 2021 R CODE"
output: word_document
---


#Question 2 Part a:
```{r}
n = 7
p = 4
X = matrix(c(rep(1,n),32,19.5,13.3,13.3,5,7.1,34.5,84.9,306.6,562,562,390.6,2175,623.5,10,9,5,5,5,3,7),n,p)
X
y = c(37.9,42.2,47.3,43.1,54.8,47.1,40.3)
b = solve(t(X)%*%X,t(X)%*%y)
b
```


```{r}
s2 = sum((y-X%*%b)^2)/(n-p)
s2
```
#Question 2 Part b:

```{r}
xst = as.vector(c(1,10,100,6))
xst %*% b + c(-1,1)*qt(0.95,df=n-p)*sqrt(s2 * t(xst) %*% solve(t(X)%*%X) %*% xst)
```

#Question 2 Part c:
```{r}
#From Slide 33 from the Test Statistic Inference for the full rank model!
C = matrix(c(0,1,0,-1),1,4)
#Calculating the Sample Standand Derivation!
s = sqrt(s2)
n = 7
#Standand error for beta1 - beta3
V = C%*%solve(t(X)%*%X)%*%t(C)%*%s2 #Our new sample variance (s2)
se = sqrt(V)
se
```


#Question 2 Part d:
```{r}
library(MASS)
C = matrix(c(0,1,0,0),1,4)
r = 1
delta = -1
Fstat = (t(C%*%b-delta)*solve(C%*%solve(t(X)%*%X)%*%t(C))*(C%*%b-delta)/r)/s2
#Fstat 
Fstat
#p value
pf(Fstat, df1= r, df2 = n-p, lower.tail = FALSE)

```


```{r}
#Attempt 1
model = lm(y~X[,-1])
summary(model)
```


#Question 2 Part e:

#Slide 61-63 IFTFRM
```{r}
SSReg = t(y) %*% X %*% b - sum(y)^2 / n
SSRes = s2*(n-p)
Fstat = (SSReg/(p-1))/(SSRes/(n-p)) 
Fstat
```

```{r}

pf(Fstat, p-1, n-p, lower.tail = FALSE)
```
#We do not reject the null hypothesis of the model relevance!


#Question 4 Part a:
```{r}
data(mtcars)
mtcars = log(mtcars[, c(1,3:7)])
pairs(mtcars,cex=0.5)
```
#COMMENT HERE: Looking at miles per gallon against the other variables, there is evidence of a linear relationship with displacement,gross horsepower,rear axle ratio,weight and a quarter mile time!

#Question 4 Part b:
```{r}
model0 = lm(mpg ~ 1, data=mtcars)
add1(model0, scope = ~.+disp+hp+drat+wt+qsec, test = "F")
```

#We take out the displacement variable
```{r}
model1 = lm(mpg ~ 1+disp, data=mtcars)
add1(model1, scope = ~.+hp+drat+wt+qsec, test = "F")
```
#Take out the weight variable

```{r}
model2 = lm(mpg ~ 1+disp+wt, data=mtcars)
add1(model2, scope = ~.+hp+drat+qsec, test = "F")
```
```{r}

#We take out horsepower
model3 = lm(mpg ~ 1+disp+wt+hp, data=mtcars)
add1(model3, scope = ~.+drat+qsec, test = "F")
```
#The final variables are disp,wt and hp!


#Question 4 Part c:
```{r}
model = step(model0, scope = ~ .+disp+hp+drat+wt+qsec)
```

#AIC is -139.83


#Question 4 Part d:
```{r}
model
```
#Were are dealing with a log transformation!
#Final model is log(mpg) = 4.8347 - 0.2553log(hp) - 0.5623log(wt) + e
#Take the exponential on both sides of the linear model
# mpg = exp(4.8347)hp^(-0.2553)wt^(-0.5623)e'
# Where e' is exp(e)

#Question 4 Part e:
```{r}
plot(model, which=1)
```
```{r}
plot(model, which=2)
```



```{r}
plot(model, which=3)
```




```{r}
plot(model, which=5)
```
#Diagnostic plots show a reasonable fit to linear model assumptions. About the only area of concern is a slight positive trend for higher fitted values and moderate leverages, but this does not appear to be too alarming.



#Question 5 Part c:
```{r}

#Using Data from Question 2
n = 7
p = 4
X = matrix(c(rep(1,n),32,19.5,13.3,13.3,5,7.1,34.5,84.9,306.6,562,562,390.6,2175,623.5,10,9,5,5,5,3,7),n,p)
y = c(37.9,42.2,47.3,43.1,54.8,47.1,40.3)

#Extracting data
X = scale(X[,-1],center=T,scale=T)
y = scale(y,center=T,scale=T)
p = 3


lambda.seq = seq(0,5,0.001)
aic.seq = c()

#generate a plot for lambda vs. AIC
for (lambda in lambda.seq){
  b = solve(t(X)%*%X + lambda*diag(p), t(X)%*%y)
  e = y - X%*%b
  SSRes = sum(e^2)
  H = X%*%solve(t(X)%*%X + lambda*diag(p))%*%t(X)
  trace = sum(diag(H))
  aic = n*log(SSRes/n)+ 2*trace
  aic.seq = c(aic.seq,aic)
}

plot(lambda.seq,aic.seq,type = "l")

```



```{r}
min(aic.seq)
#Minimum AIC
index = which.min(aic.seq)

#Lambda value
lambda.seq[index]
```









