---
title: "CSE5DEV_Lab2"
output: word_document
date: '2023-08-12'
---
#Topic 1: Data Wrangling using R.
#First we must set up the directory first, 

```{r}
setwd("C:/Users/mikel/OneDrive/Desktop/CSE5DEV Labs")

```

## Reading the data csv file.

```{r}
dat <- read.csv("data.csv")
dat

```

## 


```{r}
head(dat)
```


## Showing the column names
```{r}
names(dat)
```



## R: Conditional operators
## 1. Conditional operators for the two variables: x and y. 
```{r}
x <- 4
y <- 15
x < y
## [1] TRUE
x > y

## [1] FALSE

x<=5

## [1] TRUE

y>=20

## [1] FALSE

y == 16

## [1] FALSE

x != 5

## [1] TRUE

```
##2. Conditional operators for a vector x
```{r}
x <- c (3,5,1,2,7,6,4)
            
x < 5 # is x less than 5 

## [1] TRUE FALSE TRUE TRUE FALSE FALSE TRUE 

x <= 5 # is x less than or equal to 5

## [1] TRUE TRUE TRUE TRUE FALSE FALSE TRUE

x > 3 # is x greater than 3

## [1] FALSE TRUE FALSE FALSE TRUE TRUE TRUE

x >= 3 # is x greater than or equal to 3

## [1] TRUE TRUE FALSE FALSE TRUE TRUE TRUE 

x ==2 # is x equal to 2 

## [1] FALSE FALSE FALSE TRUE FALSE FALSE FALSE

x !=2 # is x not equal to 2

## [1] TRUE TRUE TRUE FALSE TRUE TRUE TRUE 

```

## The all and any functions check whether all or at least some entries of a logical vector are TRUE respectively.


```{r}
x <- c(3,5,1,2,7,6,4)
any(x == 2)
## [1] TRUE
all(x == 2)
## [1] FALSE
all(x < 10)
## [1] TRUE

```

## R: Logical condition operators. 
# The following example shows logical operators for a vector x

```{r}
X <- c (3 , 5 , 1 , 2 , 7 , 6 , 4 )
(x > 2 ) & (x <= 6 ) # is x greater than 2 and less than or equal to 6
## [1] TRUE TRUE FALSE FALSE FALSE TRUE TRUE
(x < 2 ) | (x > 5 ) # is x less than 2 or greater than 5
## [1] FALSE FALSE TRUE FALSE TRUE TRUE FALSE
! (x > 3 ) # not [x greater than]
## [1] TRUE FALSE TRUE TRUE FALSE FALSE FALSE
```


#Now consider this: x <- c (5, 3, 7, 9, 10).

#We want to extract the values of the vector x that are greater than 5 (7, 9, 10). There are two methods we could apply.
```{r}
#Method 1
x <- c(5, 3, 7, 9, 10)
ind <- x > 5 # is x greater than 5
print(ind)
## [1] FALSE FALSE TRUE TRUE TRUE
print(x[ind])
## [1] 7 9 10
```



```{r}

#Method 2
X <- c(5, 3, 7, 9, 10)
X[X > 5]
## [1] 7 9 10
```
#Sometimes we may need to extract data that satisfy certain criteria.

#For example, we may want to select data based on the disp value that is equal to or less than 160. We can use logical condition operators to select a subset of data.



```{r}
s <- dat[dat$disp<=160,]
print(s)
## Model mpg cyl disp hp drat wt qsec vs am gear carb
## 1 Mazda RX4 21. 0 6 160.0 110 3.90 2.620 16.46 0 1 4 4
## 2 Mazda RX4 Wag 21. 0 6 160.0 110 3.90 2.875 17.02 0 4 4
## 3 Datsun 710 22.8 4 108.0 93 3.85 2.320 18.61 1 4 1
## 8 Mere 240D 24.4 4 146.7 62 3.69 3.190 20.00 1 0 4 2
## 9 Mere 230 22.8 4 140.8 95 3.92 3.150 22.90 1 0 4 2
## 18 Fiat 128 32.4 4 78.7 66 4 .08 2 .200 19 . 47 1 4 1
## 19 Honda Civic 30.4 4 75.7 52 4.93 1.615 18 .52 1 4 2
## 20 Toyota Corolla 33.9 4 71.1 65 4.22 1.835 19.90 1 4 1
## 21 Toyota Corona 21.5 4 120.1 97 3.70 2.465 20 . 01 1 0 3 1
## 26 Fiat Xl-9 27.3 4 79.0 66 4.08 1.935 18.90 1 4 1
## 27 Porsche 914-2 26.0 4 120.3 91 4.43 2.140 16.70 0 5 2
## 28 Lotus Europa 30.4 4 95.1 113 3.77 1.513 16.90 1 5 2
## 30 Ferrari Dino 19.7 6 145.0 175 3.62 2. 770 15.50 0 5 6
## 32 Volvo 142E 21. 4 4 121.0 109 4 .11 2.780 18.60 1 4 2

```

#Alternatively, we may want to select data based on the disp value that is equal to or less than 160 AND hp less than 110.



```{r}
z <- dat[dat$disp<=160 & dat$hp<110,]
print(z)
## Model mpg cyl disp hp drat wt qsec vs am gear carb
## 3 Datsun 710 22.8 4 108.0 93 3.85 2.320 18.61 1 1 4 1
## 8 Mere 240D 24.4 4 146.7 62 3.69 3.190 20.00 1 0 4 2
## 9 Mere 230 22.8 4 140.8 95 3.92 3.150 22.90 1 0 4 2
## 18 Fiat 128 32.4 4 78.7 66 4.08 2.200 19.47 1 1 4 1
## 19 Honda Civic 30.4 4 75.7 52 4.93 1.615 18.52 1 1 4 2
## 20 Toyota Corolla 33.9 4 71.1 65 4 .22 1.835 19 . 90 1 1 4 1
## 21 Toyota Corona 21. 5 4 120.1 97 3.70 2.465 20.01 1 0 3 1
## 26 Fiat X1-9 27.3 4 79.0 66 4.08 1.935 18.90 1 1 4 1
## 27 Porsche 914-2 26.0 4 120.3 91 4.43 2.140 16 . 70 0 1 5 2
## 32 Volvo 142E 21.4 4 121.0 109 4.11 2.780 18.60 1 1 4 2

```



#Or we could select data based on the disp value that is equal to or less than 160 AND hp less than 110 for wt column.
```{r}
x <- dat[dat$disp<=160 & dat$hp<110, "Wt"]
print (x)
## [1] 2.320 3.190 3.150 2.200 1.615 1.835 2.465 1.935 2.140 2.780
z <- dat[dat$disp<=160 & dat$hp<110 ,]
print (z)
## Model mpg cyl disp hp drat wt qsec vs am gear carb
## 3 Datsun 710 22.8 4 108.0 93 3.85 2.320 18.61 1 1 4 1
## 8 Mere 240D 24.4 4 146.7 62 3.69 3.190 20.00 1 0 4 2
## 9 Mere 230 22.8 4 140.8 95 3.92 3.150 22.90 1 0 4 2
## 18 Fiat 128 32.4 4 78.7 66 4.08 2.200 19.47 1 1 4 1
## 19 Honda Civic 30.4 4 75.7 52 4.93 1.615 18.52 1 1 4 2
## 20 Toyota Corolla 33.9 4 71.1 65 4.22 1 .835 19 .90 1 1 4 1
## 21 Toyota Corona 21.5 4 120.1 97 3.70 2.465 20.01 1 0 3 1
## 26 Fiat Xl-9 27.3 4 79.0 66 4.08 1.935 18.90 1 1 4 1
## 27 Porsche 914-2 26.0 4 120.3 91 4.43 2.140 16.70 0 1 5 2
## 32 Volvo 142E 21.4 4 121.0 109 4.11 2.780 18.60 1 1 4 2

```

#R: Conditional execution 

```{r}
team_A <- 8 # Number of goals scored by Team A
team_B <- 6 # Number of goals scored by Team B
if (team_A > team_B){
print ( "Team A wins" )
}
## [1] "Team A wins"

```


#We can use multi-able else using else if as follows:
```{r}
team_A <- 4 # Number of goaLs scored by Team A
team_B <- 4# Number of goaLs scored by Team B
if (team_A > team_B){
print ( "Team A won" )
} else if (team_A < team_B){
  print ( "Team B won" )
} else {
  print("Team A & B got the same number of goals")
}

## [1] "Team A & B got same number of goals"

```



#R: Repetitive execution-for loop, repeat, while

```{r}
for (i in 1 :4 )
print (i)

## [1] 1
## [1] 2
## [1] 3
## [1] 4

i <- 1
for (i in seq ( 5 , 10 , 2))
{ print (i)}

## [1] 5
## [1] 7
## [1] 9

X <- c (2 , 1 , 3 , 6 , 1)
nr <- length (x) # vector Length
print (nr)

## [1] 5

for (i in 1 :nr)
{ print (x[i])} # print vector eLement

## [1] 2
## [1] 1
## [1] 3
## [1] 6
## [1] 1

fruit <- c ( 'Apple' , 'Orange' , 'Passion fruit' , 'Banana' )

for (p in fruit)
{
print (p)
}

## [1] "Apple"
## [1] "0range"
## [1] "Passion fruit"
## [1] "Banana"

a <- matrix (10 : 15 , 3)
print (a)

## [,1] [' 2]
## [1,] 10 13
## [2,] 11 14
## [3 ,] 12 15

nrr <- nrow(a) # number of rows is a

for (i in 1 :nrr) {print (a[i , 2])}

## [1] 13
## [1] 14
## [1] 15

```


#Consider the following example for repeat loop:

```{r}
x <- 4
repeat{
  print(x);
  x <- x + 2 ; 
  if(x > 10) {break}
  }

## [1] 4
## [1] 6
## [1] 8
## [1] 10

result <- c("Hello CSE5APG")
i <- 1
# repeat Junction
repeat {
print(result)
# update expression
i <- i + 1
# test condition
if (i > 4) {break}
}

## [1] "Hello CSE5APG"
## [1] "Hello CSE5APG"
## [l] "Hello CSE5APG"
## [l] "Hello CSE5APG"

```
#And consider the following example for while loop:

```{r}
X <- 0
while (x < 5) {print (x); x <- x + 1}
## [1] 0
## [1] 1
## [1] 2
## [1] 3
## [1] 4
i <- 1
j <- 1
mat <- matrix(0,nrow = 3,ncol= 2)
print(mat)
##         [, 1] [, 2]
## [1, ]       0     0
## [2, ]       0     0
## [3, ]       0     0


while(i<=3){
  j = 1
  while(j<=2){
    mat[i,j] <- i*2 + 1
    j <- j + 1
  }
i<-i+1
}
print(mat)
##       [, 1] [, 2]
## [1, ]     3     3
## [2, ]     5     5
## [3, ]     7     7

```



#What are R: packages and how to install R: packages
```{r}

#Ensure you installed the following packages, and open up
#in library. 
library(dplyr)
library(tidyr)
library(tidyverse)

```
#How to use R programming packages
#Step 1. Create a data frame
```{r}
df <- data.frame(color = c("blue", "black", "blue", "blue", "black"), value = 1:5)
df
```


#Step 2: Perform the following functions:
#filter()
#arrange()
#select()
#mutate()

#filter()
```{r}
filter(df,color=="blue")

```

#filter() - 2
```{r}
filter(df,value%in%c(1,4))

```
#arrange()
```{r}
arrange(df,color)

```

#arrange() - 2

```{r}
arrange(df,desc(color))

```

#select() 

```{r}
select(df,color)

```





#select() - 2

```{r}
select(df,-color)

```

#mutate()
```{r}
mutate(df,double=2*value)

```



#mutate() - 2
```{r}
mutate(df,double=2*value,quadruple = 2*double)

```


#Topic 2: Data Cleaning

#Handing Dates.
```{r}
x <- c("2015-07-01","2015-08-01","2015-09-01")
as.Date(x)
#Output
#[1] "2015-07-01" "2015-08-01" "2015-09-01"

#example using "format "
y <- c ("07/01/2017","07/01/2018","07/01/2015")
as.Date(y, format = "%m/%d/%Y")
#Output
#[1] "2017-07-01" "2018-07-01" "2015-07-01"

```

#In this example we create a date, specify the format, use a different origin and take a difference.
```{r}
# create a date
as.Date("2012-08-30")
#[1] "2012-08-30"

# specify the format
as.Date("08/30/2012",format ="%m/%d/%Y")
#[1] = "2012-08-30"

# use a different origin, for instance importing values from Excel
as.Date(41149,origin = "1900-01-01")
#[1] "2012-08-30 "

# take a difference
Sys.Date() - as.Date("1970-01-01")
#[1] Time difference of 19582 days

```

#This example prints today’s date and calculates the difference between two different dates.
#NOTE: Some of the results may change
```{r}
#print today's date
today <- Sys.Date()
format(today, format = "%B/%d/%Y")
#[1] "August/13/2023"

# alternate method with specified units
difftime(Sys.Date(), as.Date("1970-01-01"),units = "days")
#[1] Time difference of 19582 days

# see the internal integer representation
unclass(Sys.Date())
#[1] 19582

mydates <- as.Date (c("2007-06-22", "2004-02-13"))
days <- mydates[1] - mydates[2]
#[1] Time difference of 1225 days

```

#This example calculates the time between two dates using ‘difftime’ function.


```{r}
dl <- as.Date("2017-01-01")
d2 <- as.Date("2018-01-01")
print(d2 - dl)
#[1] Time difference of 365 days

difftime(d2, dl, units = "day")
#[1] Time difference of 365 days

difftime(d2, dl, units = "hour")
#[1] Time difference of 8760 hours

```


#Example: Converts the month (numeric) to month (character)

```{r}
format(today,format = "%B-%d-%Y")

```


#Handling data types
```{r}
# Conversion into Character
char1 <- as.character("hello")
#[1] "hello"
char2 <- as.character("1234")
## [1] "1234"
# Conversion into Numeric
number <- as.numeric("1234")
#[1] 1234
# Conversion into Integer
int5 <- as.integer(14.7)
#[1] 14
# Conversion into logical : 0 to False; True otherwise
logi <- as.logical(10)
#[1] True

```




```{r}
data <- data.frame (color=c(rep("blue", 5), rep ("red",
5)))
is.factor(data)
#[1] FALSE

data$color <- factor(data$color)
#[1] blue blue blue blue blue red red red red red
#[1] Levels: blue red

data$color <- as.integer(data$color)
#[1] 1 1 1 1 1 2 2 2 2 2

```


#This example shows the creation of ordered factor variables.
```{r}
ord <- c ("low", "middle", "low", "low", "low", "low",
"middle", "low", "middle",
"middle", "middle" ,"middle", "middle", "high
","high", "low", "middle",
"middle", "low", "high") 

#[1] "low" "middle" "low" "low" "low" "low"
#"middle" "low" "middle" "middle" "middle" "
#middle"  "middle" "high" "high" "low" "
#middle" "middle" "low" "high"

ord.order <- ordered(ord, levels = c("low","middle","high"))
#[1] low middle low low low low middle
#low middle middle middle middle middle high
#high low middle middle low high

#[1] Levels : low < middle < high

```



#Question: How can we handle missing values?

```{r}
Dat1<-read.csv("iris.csv")
head(Dat1)

unique(Dat1)
```

```{r}
names(Dat1)

```

```{r}
unique(Dat1)

```

```{r}
summary(Dat1)

```
#Using Dirty Iris
```{r}
Dat1<-read.csv("dirty_iris_1.csv")
head(Dat1)


```


```{r}
#Display Boolean values that contain null values
#for Petal width
is.na(Dat1$Petal.Width)

#total value of Petal Length
sum(is.na(Dat1$Petal.Length))

#total value of Petal Width
sum(is.na(Dat1$Petal.Width))

```

#Do similiar for the other ones. 
```{r}
#total value of Sepal Length
sum(is.na(Dat1$Sepal.Length))

#total value of Sepal Width
sum(is.na(Dat1$Sepal.Width))

```
#Column names in Dirty Iris 1 Dataset. 
```{r}
names(Dat1)

```


#Dimensions in Dirty Iris 1 Dataset. 
```{r}
dim(Dat1)

```


#indexing, locating where the column Petal length has null values
```{r}
ind<-which(is.na(Dat1$Petal.Length))
print(ind)

```



#
```{r}
Dat1$Petal.Length[ind]<-1.0

#One method in replacing NA's.
sum(is.na(Dat1$Petal.Length))


```


```{r}
#Here we omit from Dat1 dataset. 
Dat2 <-na.omit(Dat1)
print(Dat2)

```


```{r}
names(Dat2)
dim(Dat2)

```



```{r}
ind

```



#Topic 3: Data Normalisation

```{r}
Dat10<-read.csv("iris.csv")
head(Dat10)

```

#Apply the normalisation formula, onto the iris dataset. 

```{r}
n = function(x) (x-min(x))/(max(x)-min(x))

```

```{r}
# We take only the numerical values to normalize
iris_norm <- as.data.frame(apply(Dat10[,1:4],2,n))
iris_norm$Species <- Dat10$Species
str(iris_norm)
#Normalize the dataframe by rows
iris_norm <- as.data.frame(t(apply(Dat10[1:4],1,n)))
# Now we see that Sepal Length is always 1 because it
#is the maximum value in every row
# At the same time the Petal Width is always the lower
#value
summary (iris_norm)

```





#Refer back to Topic 1: Data Wrangling using R Page. 
#Demonstration of the gathering process

```{r}
data1 <- read.csv("data1.csv")
gather(data1,'year','cases',2:3)

```

#Demonstration of spreading process
```{r}
data2 <- read.csv("data2.csv")
gather(data2,'year','cases',2:3)

```



#Seperating and uniting processes
```{r}
data3 <- read.csv("data3.csv")
separate(data3,rate,c('cases','population'),sep="/")

```



#Demonstration of uniting process
```{r}
data4 <- read.csv("data4.csv")
unite(data4,'year',century,year,sep="")

```




```{r}
d = c(1,"a",FALSE)
class(d)

```









