---
title: "MAST30025 Week 10 Lab"
output: word_document
---

#Question 1
#Part a: Find s2 (Correct)
```{r}

#Re-typing the code from Week 8's Lab
setwd("~/Desktop/UNIMELB 2021 Material/UNIMELB S1 2021 (Currently)/MAST30025/Tutorials /Tutorials/Rfile/data")
filters = read.csv("filters.csv")

#More to add how to convert into a factor! #Slide 63 Less than full rank model!
filters$type = factor(filters$type)
```



```{r}
#Attempt 1

#y matrix
y = filters$life

#X matrix
n = dim(filters)
k = length(levels(filters$type))
X = matrix(0,n,k+1)
#Add all ones in the first column
X[,1] = 1
#Add all ones for each tao factor

#2nd column
X[filters$type==1,2] = 1
#3rd column
X[filters$type==2,3] = 1
#4th column
X[filters$type==3,4] = 1
#5th column
X[filters$type==4,5] = 1
#6th column
X[filters$type==5,6] = 1
```

```{r}
y

```

```{r}
X
```


```{r}
#Actual X matrix should works out exactly above
X1 = matrix(0,30,6)
X1[,1] = 1
for (i in 1:5){X1[filters$type==i,i+1] = 1}
X1
```

```{r}
#Actual X matrix should works out exactly above
t(X)%*%X
```

```{r}
#If you want it to convert into a matrix form!
#finding the inverse
XtXc = matrix(0,6,6)
XtXc[2:6,2:6] = solve((t(X)%*%X)[2:6,2:6])
XtXc
```
```{r}
#Finding another conditional inverse (Open the MASS Package)
XtXc2 = ginv(t(X)%*%X)
XtXc2
```

```{r}
#Finding the least squares estimator 
b = XtXc2%*%t(X)%*%y
b

```
```{r}
#Finding s2
e = y - X%*%b
SSRes = sum(e^2)
s2 = SSRes/(30-5)
s2
```



#Part b: 95% Confidence Interval for the difference in lifespan between filter types 3 and 4.(correct) 
```{r}
tt = c(0,0,0,1,-1,0)
ta = qt(0.975,30-5)
halfwidth = ta*sqrt(s2*t(tt)%*%XtXc2%*%tt)
c(tt%*%b-halfwidth,tt%*%b+halfwidth)

```

#Part c: Show that the hypothesis that the filters all have the same lifespan is testable. (Nearly there!)
```{r}
#Using theorm 6.10
#Type 1 and 2
tt1 = c(0,1,-1,0,0,0)
round(tt1%*%XtXc2%*%t(X)%*%X)

```


```{r}
#Using theorm 6.10
#Type 1 and 3
tt2 = c(0,1,0,-1,0,0)
round(tt2%*%XtXc2%*%t(X)%*%X)

```




```{r}
#Using theorm 6.10
#Type 1 and 4
tt3 = c(0,1,0,0,-1,0)
round(tt3%*%XtXc%*%t(X)%*%X)

```

```{r}
#Using theorm 6.10
#Type 1 and 5
tt4 = c(0,1,0,0,0,-1)
round(tt%*%XtXc2%*%t(X)%*%X)

```

```{r}
#Using theorm 6.10
#Type 2 and 3
tt5 = c(0,0,1,-1,0,0)
round(tt5%*%XtXc2%*%t(X)%*%X)

```

```{r}
#Using theorm 6.10
#Type 2 and 4
tt6 = c(0,0,1,0,-1,0)
round(tt6%*%XtXc2%*%t(X)%*%X)

```




```{r}
#Using theorm 6.10
#Type 2 and 5
tt7 = c(0,0,1,0,0,-1)
round(tt7%*%XtXc2%*%t(X)%*%X)

```


```{r}
#Using theorm 6.10
#Type 3 and 4
tt8 = c(0,0,0,1,-1,0)
round(tt8%*%XtXc2%*%t(X)%*%X)

```



```{r}
#Using theorm 6.10
#Type 3 and 5
tt9 = c(0,0,0,1,0,-1)
round(tt9%*%XtXc2%*%t(X)%*%X)

```

```{r}
#Using theorm 6.10
#Type 4 and 5
tt10 = c(0,0,0,0,1,-1)
round(tt10%*%XtXc2%*%t(X)%*%X)

```


#Suprising they are all testable!
#For all having the same lifespan that is testable!


#Actual Solution!
```{r}
#Using theorm 6.10
C = matrix(c(0,1,-1,0,0,0,0,0,1,-1,0,0,0,0,0,1,-1,0,0,0,0,0,1,-1),4,6,byrow=TRUE)
round(C%*%XtXc2%*%t(X)%*%X)
#Where C is at full rank! From Slide 6 Inference for the less than full rank model!
```


#Part d: Test this hypothesis, using matrix theory! 
```{r}
#Attempt 1
numer = t(C%*%b)%*%solve(C%*%XtXc%*%t(C))%*%C%*%b/4 #Missing Bracket from
#ginv(t(X)%*%X)
#Theorem 7.3 derived from the lectures, it is important!
Fstat = numer/s2
pf(Fstat,4,25,lower=F)

```
#We can reject the null hypothesis that all filters have the same lifespan.


#Part e: Test the same hypothesis using the linearHypothesis function from the car package. (ASK ABOUT THE C2 MATRIX)
```{r}
library(car)
model = lm(filters$life~filters$type,data = filters)
C2 = matrix(c(0,1,0,0,0,0,0,1,0,0,0,0,0,1,0,0,0,0,0,1),4,5,byrow=TRUE)
linearHypothesis(model,C2,rep(0,4))
```


#Question 4
#Part a: #Module 7 slide 23 reading from the Tennis ball example!
#Test the hypothesis that the work plan has no effect on the absenteeism. 
```{r}
#Attempt 1
n = 275
r = 3
XtXc = diag(c(0,1/c(100,85,90)))
b = c(0,9,6.2,10.1)
s2 = 110.15
C = matrix(c(0,1,-1,0,0,0,1,-1),2,4,byrow = T) #Depends on your C matrix also add byrow = T
Fstat = (t(C%*%b)%*%solve(C%*%XtXc%*%t(C))%*%C%*%b/2/s2) #Theorem 7.3 understand your rank of hypothesis!
pf(Fstat,2,n-r,lower=F)
```
#We reject the null hypothesis, conclude that work plan has no effect on the absenteeism. 




#Part b: Test the hypothesis that work plans 1 and 3 have the same rate of absenteeism
```{r}
#Attempt 1
n = 275
r = 3
XtXc = diag(c(0,1/c(100,85,90)))
b = c(0,9,6.2,10.1)
s2 = 110.15
C1 = matrix(c(0,1,0,-1),1,4,byrow = T) #Depends on your C matrix
Fstat = (t(C1%*%b)%*%solve(C1%*%XtXc%*%t(C1))%*%C1%*%b/s2)
pf(Fstat,1,n-r,lower=F)
```
#We do not reject the null hypothesis, conclude that work plans 1 and 3 do not have the same rate of absenteeism. 








