---
title: "MAST30025 Assignment 1 S1 2020"
output: word_document
---
## Question 3

#Part a: Describe the distribution of Ay.
```{r}

A = matrix(c(2,0,-1,0,3,0,-1,0,2)/3,3,3)
V = matrix(c(2,0,1,0,1,0,1,0,2),3,3)
mu = t(matrix(c(1,-1,0),1,3))
```

## 


#Mean of the Ay distribution 
```{r}
A%*%mu
```

#Variance of the Ay distribution
```{r}
A%*%V%*%t(A)

```




## Question 3

#Part b: Find E[t(y)Ay].

#Using Theorem 3.2
```{r}
sum(diag(A%*%V)) + t(mu)%*%A%*%mu
```


## Question 3

#Part c: Describe the distribution of t(y)%*%A%*%y
```{r}
lamba = 0.5%*%t(mu)%*%A%*%mu
lamba  #Our noncentrality parameter 
```




## Question 5

#Part a: 
```{r}
y = matrix(c(8.5,8,7.5,10,11,15,13.5,14),8,1)
y
```


```{r}
X = cbind(rep(1,8),c(1.35,1.33,2,1.4,1.43,1.2,1.3,1.28),c(50,55,60,52,47,45,49,63))
X
```

# Part b: Least square estimators 
```{r}
b = solve(t(X)%*%X, t(X)%*%y)
b #b0,b1 and b2 from each row respectively 
```

#Part c: Calculate the residual sum of squares SSres and sample variance s^2. 
```{r}
e = y - X%*%b 
e

```

#Now t o find s2
```{r}
N=8
P=3
SSRES = sum(e^2)
s2 = SSRES/(N-P)
s2
```

#Part d: Predict (using a point estimate) the average fuel mileage of a car which weighs 1.8 tons and is driven at 59 km/hr.
```{r}
c(1,1.8,59)%*%b


```




# Part b: Least square estimators 
```{r}

t(X)%*%X


```


```{r}

t(X)%*%y


```




