---
title: "Assignment 1 S1 2019 Question 4 and 5"
output: word_document
---




#Question 4 
#Part a: Describe the distribution of Ay
```{r}
A = matrix(c(4,-2,0,-2,1,0,0,0,10)/10,3,3)
A
```


#Ay ~ MVN(A*mu,AVt(A)) 
```{r}
mu = matrix(c(3,0,-2),3,1)
mu

```


#mean distribution A%*%mu
```{r}
A%*%mu
 
```
#variance distribution A%*%V%*%t(A)
```{r}
V = matrix(c(2,0,0,0,2,0,0,0,1),3,3)
A%*%V%*%t(A)
 
```





#Part b:Find E[t(y)Ay]
```{r}
sum(diag(A%*%V)) + t(mu)%*%A%*%mu
 
```
#Part c:Describe the distribution of t(y)%*%A%*%y
```{r}
rankMatrix(A)[1] 
 
```

#AV is idempotent has a noncentral chi squared distribution with 2 degrees of freedom
# and has a non centrality parameter.

```{r}
lamba = 0.5*t(mu)%*%A%*%mu
lamba
 
```

#Part d: Find all linear combinations of y elements which are independent of t(y)%*%A%*%y
```{r}
A 
 
```

```{r}
rref(A) #install pracma package 
 
```

columns 1 and 2 are mutiples of each other c1 + 2c2 = 0 and c3 = 0



#Question 5: 
#Part a: The linear model is of the form y = X*beta + e (where e is epilson). Write down the matrices and vectors involved in this equation. 
```{r}
y = matrix(c(27.3,42.4,38.7,4.5,23,166.3,109.7,80.1,150.7,20.3,189.7,131.3,404.2,149),14,1)
y
```



```{r}
X = matrix(c(rep(1,14),13.1,15.3,25.8,1.8,4.9,55.4,39.3,26.7,47.5,6.6,94.7,61.1,135.6,47.6),14,2)
X
```


#Part b: Find the least squares estimates of the parameters 
```{r}
b = solve(t(X)%*%X, t(X)%*%y)
b
```


#b0 and b1 respectively for each row 




#Part c: Calculate the sample variance s2.
```{r}
e = y - X%*%b
e #Residual errors

```


```{r}
n = 14 #sample size
p = 2 #number of parameters
SSRes = sum(e^2)
ssquared = SSRes/(n-p)
ssquared


```


#Part d: A fisherman sold ocean trout for 18 cents per round in 1970. Predict the price for ocean trout in 1980.
```{r}
c(1,18)%*%b
```


#Part e: Calculate the standardised residual for the sea scallops.
```{r}
X

```


```{r}
a = solve(t(X)%*%X)
a
```

```{r}
H = X%*%a%*%t(X)
H

```


```{r}
z = e/sqrt(ssquared * (1 - diag(H)))
z[13]
```




#Part f: Calculate the Cook’s distance for sea scallops.
```{r}
k = 1
D = z^2 * (diag(H)/(1-diag(H))) * 1/(k+1)
D[13]
```
#Part g:Does sea scallops fit the linear model? Justify your argument

```{r}
plot(X[,2],y)
abline(b[1],b[2])

```

Full explaination: The Cook’s distance certainly indicates it should be of some concern; however looking at the plot, it seems that the fit is actually okay. There is considerable evidence for heteroskedasticity — the variance increases with x (the design variable). Sea scallops has (by far) the largest x and so may be prone to a larger variance than the remaining points. The high Cook’s distance therefore comes primarily from a very high leverage, rather than a bad fit to the model.




