---
title: "MAT4MDSTutorial5"
output: word_document
date: '2023-02-22'
---

#Question 1 Part a.
#Creating the dataframe for ice core samples 

```{r}
#Depth (in meters)
Depth_m = c(127,303,376,474,602,748,853,976)

#CO2 (ppm)
CO2_ppm = c(275,259,245,195,223,179,201,201)

df = data.frame(Depth_m,CO2_ppm)
df
```

#Now to estimate the linear least squares line of best fit for estimating CO2 level from depth.
```{r}
xval = df$Depth_m
xval
```


```{r}
yval = df$CO2_ppm
yval
```
### Linear least squares Regression: Matrix Method
### Refer using LLSR using R

### Getting the values for the coffients Matrix A
# evaluate sum of the x-values
```{r}
x_total = sum(xval)
x_total
```

# evaluate sum of the squared x-values
```{r}
x_sq_total = sum(xval^2)
x_sq_total
```
# set n as number of entries in the vector 'xval'.
```{r}
n = length(xval)
n

```

### get the values for the constants Matrix B. 

# evaluate sum of the y-values.
```{r}
y_total = sum(yval)
y_total
```


# evaluate sum of x-values*y-values.
```{r}
prod_x_y = sum(xval*yval)
prod_x_y
```

# Set the coefficients matrix
```{r}
A = matrix(c(x_sq_total,x_total,x_total,n),2,2)
A
```


# Set the constants matrix

```{r}
B = matrix(c(prod_x_y,y_total))
B
```


# Solve the system AX=B, where X = (alpha,beta) from the linear regression
# yval = alpha*xval + beta

```{r}
X = solve(A,B)
X
```

## My answer: yval =  -0.09698511*xval + 276.30707791


#Question 1b. Use your answer to (a) to estimate the CO2 concentration at a depth of 1050m. 
```{r}
x = 1050
y_1050m = -0.09698511*x + 276.30707791
y_1050m
```

#See the plot for visualisation

```{r}
linmod  = lm(yval~xval, data=df)
linmod
```

# plot dotplot (scatter plot) for the original points: xval, yval

```{r}
plot(xval,yval,xlab = 'Depth in meters',ylab = ' CO2 in ppm (parts per million)')
abline(linmod,col='red')

```



#Question 1c. Refering to the alpha value, According the our model we 
# gradually decrease the level concentration CO2
# as we increase the depth during the last 60,000 years. 


#Question 2.
#Part a. 
#Estimate the depth at which an ice core sample was obtained if
# is known that its CO2 concentration is 200 ppm. 
```{r}
y = 200  
x = (y - 276.30707791)/(-0.09698511)
x
```
#Part b. Explain carefully why you needed a different line of best fit in Question 2a from the one derived in Question 1.

```{r}
summary(linmod)
```

#Because it determines whether the model describes 
#how good the model can properly fit. 
#How many number of data points can it pass through.

#It will require to modify between the dependent 
# variables and independent variables. 



#Question 4
#Part a
```{r}
#Snow cover (10^6*km^2)
Snowc = c(6.6,7.7,8.1,18.2,26.6,29.4)

#Wind Stress (N/m2)
Ws= c(.125,.155,.196,.106,.062,.024)

df = data.frame(Snowc,Ws)
df
```



```{r}
xval = df$Snowc
yval = df$Ws
```


```{r}
x_total = sum(xval)
x_total
```

```{r}
x_sq_total = sum(xval^2)
x_sq_total
```

```{r}
n = length(xval)
n
```

```{r}
y_total = sum(yval)
y_total
```
```{r}
prod_x_y = sum(xval*yval)
prod_x_y
```

```{r}
A = matrix(c(x_sq_total,x_total,x_total,n),2,2)
A
```

```{r}
B = matrix(c(prod_x_y,y_total))
B
```



```{r}
X = solve(A,B)
X
```


#yval = -0.005547874*xval+0.200654098


#Question 4b.
#What does this model reveal about the nature and strength of the effect of decreasing snow cover on wind stress?

#The model reveals the prediction for the best estimate for the effect of decreasing snow cover on wind stress.


#Question 5a. 

```{r}
#Year (Year-1951)
year = c(0,10,20,30,40,50,60)

#Population (in millions)
pop = c(361,439,548,683,846,1029,1211)

df = data.frame(year,pop)
df
```




```{r}
xval = df$year
yval = df$pop
```



```{r}
x_total = sum(xval)
x_total
```


```{r}
x_sq_total = sum(xval^2)
x_sq_total
```


```{r}
n = length(xval)
n
```
```{r}
y_total = sum(yval)
y_total
```

```{r}
prod_x_y = sum(xval*yval)
prod_x_y
```

```{r}
A = matrix(c(x_sq_total,x_total,x_total,n),2,2)
A
```



```{r}
B = matrix(c(prod_x_y,y_total))
B
```


```{r}
X = solve(A,B)
X
```

#Substitute into the exponential model. 
#yval = 299.42857*exp(14.38571*xval)

#Take the natural logarithm of the exponential model. 

#Final model. 
#yval = log(299.42857)+14.38571*xval

#NOTE: log() in R takes the function of natural logarithm (base e).

#Question 5b.  According to this model, what is the estimate of the population in 2011? Compare that value to the actual population. [In this case, the discrepancy is due to the lower growth rate from 2001 to 2011 than in the period 1951 to 2001.

```{r}
xval = 2011-1951
yval = log(299.42857)+14.38571*xval
yval

```

#868.8445 in millions people (0.868 billion people)
#The actual value is (1.211 billion people)

#Teacher can have further comment here.






